Randomforest

### 2019-05-13 partialrda.Rmd
Minor
Genus and core

```{r libraries, cache = FALSE}
library(dplyr)
library(vegan)
library(broom)
library(tibble)
library(knitr)
library(phyloseq)
library(microbiome)
```

@@

```{r functions}
myresults.table <- function(model) {
    data.frame(Inertia = c(
                   Conditioned = summary(model)$partial.chi,
                   Constrained = summary(model)$constr.chi,
                   Unconstrained = summary(model)$unconst.chi)) %>%
        tibble::rownames_to_column("Variance") %>%
        mutate(Proportion = Inertia/sum(Inertia)) %>%
        mutate_at(2:2, funs(round(., 1))) %>%
        mutate_at(3:3, funs(round(., 5))) %>%
        kable
}

anotab <- function(x) {
    EPS <- sqrt(.Machine$double.eps)
    Pval <- (colSums(sweep(x$F.perm, 2, x$F.0 - EPS, ">=")) + 1)/(x$nperm + 1)
    anotab <- data.frame(x$df, x$chi, c(x$F.0, NA), c(Pval, NA))
    colnames(anotab) <- c("Df", "Inertia", "F", "Pr(>F)")
    rownames(anotab) <- c(x$termlabels, "Residual")
    class(anotab) <- c("anova", "data.frame")
    anotab 
}
```

@@

```{r models}
rda.species <- rda(rda.abundances.species ~ MAP +
                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                   data = meta(pseq.species))
rda.genus <- rda(rda.abundances.genus ~ MAP +
                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                 data = meta(pseq.genus))
rda.genus.core <- rda(rda.abundances.genus.core ~ MAP +
                       Condition(BL_AGE + SEX + BMI + CURR_SMOKE + Q57X + DIAB +
                       BL_USE_RX_C03 + BL_USE_RX_C07 + BL_USE_RX_C08 + BL_USE_RX_C09),
                   data = meta(pseq.genus.core))

```

```{r myresults species}
myresults.table(rda.species)
```

```{r myresults genus}
myresults.table(rda.genus)
```

```{r myresults genus core}
myresults.table(rda.genus.core)
```

# ANOVAs

Species

```{r anova species}
anova.species <- permutest(rda.species, method="reduced", perm=1)
anova.species %>% anotab %>% kable
```

Genus

```{r anova genus}
anova.genus <- permutest(rda.genus, method="reduced", perm=1)
anova.genus %>% anotab %>% kable
```

```{r anova genus core}
anova.genus.core <- permutest(rda.genus.core, method="reduced", perm=1)
anova.genus.core %>% anotab %>% kable
```

# Plots

Species

```{r rda plot species}
plot(rda.species)
```

Genus

```{r rda plot species}
plot(rda.genus)
```

Genus core

```{r rda plot species}
plot(rda.genus.core)
```

>>>>

```{r libraries, cache = FALSE}
library(dplyr)
library(vegan)
library(broom)
library(tibble)
library(knitr)
library(phyloseq)
library(microbiome)
library(purrr)
library(parallel)
```

@@

@@

```{r model function}
modelfunction <- function(x, rda.abund, df.meta) {
    covariates <- c("BL_AGE", "SEX", "BMI", "CURR_SMOKE", "Q57X",
                   "DIAB", "BL_USE_RX_C03", "BL_USE_RX_C07",
                   "BL_USE_RX_C08", "BL_USE_RX_C09")
    fo <- sprintf("rda.abund ~ %s + Condition(%s)", x, paste0(covariates, collapse = "+"))
    model <- rda(as.formula(fo), data = df.meta)
    anova.term <- anova(model, by="term", perm=100)
    data.frame(var = as.character(x),
               beta = coef(model)[,1][[ifelse(x == "HYPERTENSION", "HYPERTENSION1", x)]],
               RDA.Conditioned = summary(model)$partial.chi,
               RDA.Constrained = summary(model)$constr.chi,
               RDA.Unconstrained = summary(model)$unconst.chi,
               p.anova = anova.term[["Pr(>F)"]][1],
               fo = fo)
}
```

```{r return parser}
returnparser <- function(ret){
    res <- do.call(rbind, ret) %>%
        mutate(RDA.total =  RDA.Conditioned + RDA.Constrained + RDA.Unconstrained,
               RDA.Conditioned.percent = RDA.Conditioned/RDA.total,
               RDA.Constrained.percent = RDA.Constrained/RDA.total,
               RDA.Unconstrained.percent = RDA.Unconstrained/RDA.total
               )

    res %>%
        dplyr::select(var,
                      beta,
                      p.anova,
                      RDA.Conditioned.percent,
                      RDA.Constrained.percent,
                      RDA.Unconstrained.percent) %>%
        kable()
}
```

## Species

```{r models species}
ret.species <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.species,
              df.meta = meta(pseq.species),
              mc.cores = 5)
returnparser(ret.species)
```

## Genus

```{r models genus}
ret.genus <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.genus,
              df.meta = meta(pseq.genus),
              mc.cores = 5)
returnparser(ret.genus)
```

## Genus core

```{r models genus core}
ret.genus.core <- mclapply(c2l("SYSTM", "DIASM", "PULSEPRESSURE", "MAP", "HYPERTENSION"),
              modelfunction,
              rda.abund = rda.abundances.genus.core,
              df.meta = meta(pseq.genus.core),
              mc.cores = 5)
returnparser(ret.genus.core)
```
